---
title: "projectAndCluster"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r, include =FALSE}
source("scripts.R")
```

## Load data 

```{r}
counts <- readMM("countsMatrix.mtx")
genes <- read.csv("genes.csv", header = F)$V1
cells<- read.csv("cells.csv", header = F)$V1
rownames(counts) <- genes
colnames(counts) <- cells
meta <- as.data.frame(read.csv("meta.csv"))
labels <- as.data.frame(read.csv("labels.csv"))
```

## Normalize and select HVGs

We then normalize the data by library size, and select HVGs for projection and clustering based on mean expresion, and fano factor.

```{r}
meta$umi <- Matrix::colSums(counts)
countn <- normalize(counts)
hvg <- compute_hvg(countn, fano = 0.65)
```

## Projection

We then project the data to two dimensions using UMAP. (Note, slightly different projections will be generated based on the choice of seed.)

```{r}
meta <- umap_project(meta,hvg, neighbours = 20, dist = 0.5, seed = 12)
plotMeta(meta)
```


## Clustering

We then cluster the data by combining the walktrap graphical clustering algorithm with the fuzzy simplicial set from UMAP. (Note, changing the versions of the python modules, e.g. umap, can alter the resolution of the clustering. See the environment.yaml file for the versions used when creating this notebook.)

```{r}
meta <- umap_cluster(meta,hvg, neighbours = 10, steps = 10, seed1 = 3, seed2 = 8)
plotMeta(meta, mode = "cluster")
```



```{r}
sessionInfo()
```
